---
title: "Additional Insights"
output: html_notebook
---



Model Matrix:

```{r}
model.matrix(Percent_Bachelors~Rural_urban_code,data=full_county_table)
```



Linear Regression Models:

```{r lin_models}
income_model <- lm(Percent_Bachelors~Median_Household_Income_2019,data=full_county_table)
summary(income_model)

growth_model <- lm(Percent_Bachelors~avg_decade_growth_rate,data=full_county_table)
summary(growth_model)

poverty_model <- lm(Percent_Bachelors~Percent_Poverty,data=full_county_table)
summary(poverty_model)


```

The greatest single indicator of county-level education that we examined is median income. According to the linear model I created, median income accounts for 50.76% of the variation among counties in percentage of people receiving a bachelors degree. 


```{r lin_models_diploma}
income_no_diploma_model <- lm(Median_Household_Income_2019~Percent_No_Diploma,data=inc_edu_county)
summary(income_no_diploma_model)

percent_bach_no_diploma_model <- lm(Percent_No_Diploma~Percent_Bachelors,data=inc_edu_county)
summary(percent_bach_no_diploma_model)

percent_bach_no_diploma_model <- lm(Percent_No_Diploma~Percent_Bachelors,data=inc_edu_county)
summary(percent_bach_no_diploma_model)
```


Multiple Linear Regression Models:

```{r multi_models}
multi_model <- lm(Percent_Bachelors~Median_Household_Income_2019+avg_decade_growth_rate+Percent_Poverty,data=full_county_table)
summary(multi_model)

```

```{r density_plot_bach}
densityPlot(inc_edu_county$Percent_Bachelors, "Density Plot", "Percent Bachelors")
```
  
County Table Grouped By Urban Classification

```{r group_by_urb}
bach_urban_class <- full_county_table %>%
group_by(Rural_urban_code)%>%
  summarise_at(vars(Percent_Bachelors),funs(mean(Percent_Bachelors)))

no_diploma_urban_class <- full_county_table %>%
group_by(Rural_urban_code)%>%
  summarise_at(vars(Percent_No_Diploma),funs(mean(Percent_No_Diploma)))

diploma_urban_class <- full_county_table %>%
group_by(Rural_urban_code)%>%
  summarise_at(vars(Percent_Diploma),funs(mean(Percent_Diploma)))

assoc_urban_class <- full_county_table %>%
group_by(Rural_urban_code)%>%
  summarise_at(vars(Percent_Associates),funs(mean(Percent_Associates)))

```

#Graphs:

```{r graphs}

bach_point(full_county_table$Median_Household_Income_2019)
bach_smooth(full_county_table$Median_Household_Income_2019)

bach_point(full_county_table$avg_decade_growth_rate)
bach_smooth(full_county_table$avg_decade_growth_rate)

bach_point(full_county_table$Percent_Poverty)
bach_smooth(full_county_table$Percent_Poverty)

bach_boxplot(full_county_table$Rural_urban_code)

```

```{r functions}
densityPlot = function(x, plot_name, x_name){
  x <- na.omit(x)
  hist(x_var, col = "red", border = "black", prob = TRUE, main = plot_name, xlab = x_name)
  lines(density(x_var), col = "blue", lwd = 2)
}

bach_point = function(x_var){
  ggplot(full_county_table, aes(x_var, Percent_Bachelors))+
  geom_point()
}

bach_smooth = function(x_var){
  ggplot(full_county_table, aes(x_var, Percent_Bachelors))+
  geom_smooth()
}

bach_boxplot = function(x_var){
  ggplot(full_county_table, aes(x_var, Percent_Bachelors))+
  geom_boxplot()
}

point_func = function(x,y,x_name,y_name){
  ggplot(full_county_table, aes(x_var, Percent_Bachelors))+
  geom_point()+
  lab(x_name,y_name)
}

urb_col <- function(y_var){
  ggplot(grouped_by_urban_class, aes(Rural_urban_code,y_var,fill = Rural_urban_code))+
  geom_col()+
  theme(legend.position = "none")
}
```

```{r group_by_urb_graphs}
ggplot(group_by_urb_table, aes(Rural_urban_code,Percent_No_Diploma,fill = Rural_urban_code))+
  geom_col()+
  theme(legend.position = "none")

ggplot(group_by_urb_table, aes(Rural_urban_code,Percent_Diploma,fill = Rural_urban_code))+
  geom_col()+
  theme(legend.position = "none")

ggplot(group_by_urb_table, aes(Rural_urban_code,Percent_Associates,fill = Rural_urban_code))+
  geom_col()+
  theme(legend.position = "none")

ggplot(group_by_urb_table, aes(Rural_urban_code,Percent_Bachelors,fill = Rural_urban_code))+
  geom_col()+
  theme(legend.position = "none")
```

(Part of) Old Summary:
  organized the dataframe according to our preference to establish an relationship.
          We have grouped the findings as per state and counties in the United States of America and visualized it through the use of plot_usmap function.
          This plot shows the distribution percentage of student enrollment to different programs in a county or state.
          Furthermore we have created correlation plot to identify the connection between income and programs enrolled, it is also made dynamic to make it easier to understand the link between two variables.
          In addition to it, we have performed linear model to get a deeper knowledge on the relation.
          We have put together all our findings in one place with the help of 'Shiny'.
          We have used shiny dashboard to summarize our thoughts and process.")